---
title: "SidewalkShed"
format: html
editor: visual
---

## Sidewalk Sheds

```{r}
library(HelpersMG)
library(readr)
library(tidyverse)
library(caret)
library(mgcv)
library(sf)
library(tidycensus)
library(DHARMa)
library(httr) 
library(ggplot2)
library(viridis)


# Read currently active permits file
D_AP <- read.csv("Active_Sheds_0208.csv")

# process active permits data
D_AP_2 <- D_AP %>%
  mutate(
    active_created_parsed = parse_date_time(D_AP$First.Permit.Date, orders = c('mdy', 'ymd')),
    active_year = year(ymd(active_created_parsed)),
    active_expired_parsed = parse_date_time(D_AP$Permit.Expiration.Date, orders = c('mdy', 'ymd')),
    active_expire_year = year(ymd(active_expired_parsed))
  ) %>%
  rename( job__ = Job.Number) %>% 
  filter(
    is.na(Longitude.Point) == FALSE,
    is.na(Latitude.Point) == FALSE,
  ) %>%
    st_as_sf(coords = c("Longitude.Point", "Latitude.Point")) %>%
    st_set_crs(4326)

```

```{r}
head(D_AP)
unique(D_AP$activity)
```

```{r}
D_AP_census_tract[D_AP_census_tract$NAME == "Census Tract 530, Queens County, New York", ]
```

```{r}
# Get Census Data
my_counties <- c("New York County", "Bronx County", "Queens County", "Kings County", "Richmond County")

D_ACS <- get_acs(
  geography = "tract",
  state = "NY",
  county = my_counties,
  variables = c(
    med_inc = "B19013_001",
    white = "B02001_002",
    black = "B02001_003",
    poverty = "B17001_002",
    tenure = "B25003_001",
    black_tenure = "B25003B_001",
    per_tenure_size = "B25010_001",
    med_bdr_rent = "B25031_001",
    med_year_build = "B25035_001",
    total_building = "B25034_001",
    age_2014 = "B25034_002",
    age_2010 = "B25034_003",
    age_2000 = "B25034_004",
    med_gross_rent = "B25064_001",
    med_value = "B25077_001",
    rent_by_income = "B25106_001"
  ),
  summary_var = "B01003_001",
    geometry = TRUE
) %>%
   select(-moe, -summary_moe) %>%
  pivot_wider(names_from = "variable", values_from = "estimate") %>%
    st_as_sf() %>%
    mutate(
        black_perc = black / summary_est,
        pov_perc = poverty / summary_est,
        age_before_2000 = total_building - age_2014 - age_2010 - age_2000
    ) %>%
  st_set_crs(4326)

```


The `echo: false` option disables the printing of code (only output is displayed).

```{r}
D_AP_acs <- D_AP_2 %>% st_join(D_ACS, left = FALSE) 

D_AP_census_tract <- D_AP_acs %>% 
  data.frame() %>%
  group_by(NAME) %>%
  summarise(
    geometry = geometry[1],
    n_active = n(),
    total_age = sum(Age),
    avg_age = total_age/n_active,
    long = avg_age > 513,
    prevalent = n_active > 5.5,
    black_perc = black_perc[1],
    pov_perc = pov_perc[1],
    med_inc = med_inc[1],
    summary_est = summary_est[1],
    tenure = tenure[1],
    black_tenure = black_tenure[1],
    per_tenure_size = per_tenure_size[1],
    total_building = total_building[1],
    age_before_2000 = age_before_2000[1],
    med_bdr_rent = med_bdr_rent[1],
    med_year_build = med_year_build[1],
    med_gross_rent = med_gross_rent[1],
    med_value = med_value[1],
    rent_by_income = rent_by_income[1],
  ) %>%
  filter(age_before_2000 >= 0) %>%
  st_as_sf() %>%
    na.omit()

D_AP_census_tract_11 <- D_AP_acs %>% 
  data.frame() %>%
  filter(activity == "Local Law 11") %>%
  group_by(NAME) %>%
  summarise(
    geometry = geometry[1],
    n_active = n(),
    total_age = sum(Age),
    avg_age = total_age/n_active,
    long = avg_age > 513,
    prevalent = n_active > 5.5,
    black_perc = black_perc[1],
    pov_perc = pov_perc[1],
    med_inc = med_inc[1],
    summary_est = summary_est[1],
    tenure = tenure[1],
    black_tenure = black_tenure[1],
    per_tenure_size = per_tenure_size[1],
    total_building = total_building[1],
    age_before_2000 = age_before_2000[1],
    med_bdr_rent = med_bdr_rent[1],
    med_year_build = med_year_build[1],
    med_gross_rent = med_gross_rent[1],
    med_value = med_value[1],
    rent_by_income = rent_by_income[1],
  ) %>%
  filter(age_before_2000 >= 0) %>%
  st_as_sf() %>%
    na.omit()


D_AP_census_tract_C <- D_AP_acs %>% 
  data.frame() %>%
  filter(activity != "Local Law 11") %>%
  group_by(NAME) %>%
  summarise(
    geometry = geometry[1],
    n_active = n(),
    total_age = sum(Age),
    avg_age = total_age/n_active,
    long = avg_age > 513,
    prevalent = n_active > 5.5,
    black_perc = black_perc[1],
    pov_perc = pov_perc[1],
    med_inc = med_inc[1],
    summary_est = summary_est[1],
    tenure = tenure[1],
    black_tenure = black_tenure[1],
    per_tenure_size = per_tenure_size[1],
    total_building = total_building[1],
    age_before_2000 = age_before_2000[1],
    med_bdr_rent = med_bdr_rent[1],
    med_year_build = med_year_build[1],
    med_gross_rent = med_gross_rent[1],
    med_value = med_value[1],
    rent_by_income = rent_by_income[1],
  ) %>%
  filter(age_before_2000 >= 0) %>%
  st_as_sf() %>%
    na.omit()
```

### EDAs

#### Where is the oldest sidewalk sheds?

```{r}
D_AP_2[order(-D_AP_2$Age),]
```
#### In long-lasting sidewalk sheds, are there more commercial or residential buildings?

```{r}
D_AP_type <- D_AP_2 %>% filter(Age > 513) %>% group_by(Commercial) %>% 
  summarise (count = n(),
             average_Age = sum(Age) / count)

ggplot(D_AP_type, aes(x = Commercial, y = count, fill = Commercial)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = count), vjust = -0.3) +
  labs(title="Among long-standing sidewalk sheds (Age>513), there are more non-commercial sheds")
```


```{r}
D_AP_activity <- D_AP_2 %>% group_by(activity) %>% 
  summarise (count = n(),
             total_Age = sum(Age),
             average_Age = sum(Age) / count)

head(D_AP_activity)

D_AP_activity_long <- pivot_longer(D_AP_activity, cols = c(count, average_Age), names_to = "variable", values_to = "value")

ggplot(D_AP_activity, aes(x = activity, y= count, fill = activity)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.3) +
  labs(title="Among long-standing sidewalk sheds (Age>513), there are more non-commercial sheds")

ggplot(D_AP_activity, aes(x = activity, y= average_Age, fill = activity)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = average_Age), vjust = -0.3) +
  labs(title="Among long-standing sidewalk sheds (Age>513), there are more non-commercial sheds")

```

```{r}
D_AP_census_tract_11 %>%
      st_as_sf() %>%
    ggplot(aes(color = n_active, fill = n_active)) +
    geom_sf()

```
```{r}
colnames(D_AP_census_tract)
```


```{r}
D_AP_census_tract_2 <- D_AP_census_tract %>%
      st_as_sf() %>%
    filter(avg_age < 4000) 
    
ggplot(D_AP_census_tract_2, aes(color = avg_age, fill = avg_age)) +
    geom_sf()
  
ggplot(D_AP_census_tract, aes(color = long, fill = long)) +
  geom_sf() +
  viridis::scale_fill_viridis(discrete = TRUE,direction=-1) + 
  labs(title = "Census tract with sidewalk sheds ages longer than averages")

ggplot(D_AP_census_tract, aes(color = n_active, fill = n_active)) +
    geom_sf() + 
  labs(title = "Fewer Sidewalk Sheds Permits in Poorer Areas")

ggplot(D_AP_census_tract, aes(color = pov_perc, fill = pov_perc)) + geom_sf() +
  labs(title = "Fewer Sidewalk Sheds Permits in Poorer Areas 2")

ggplot(D_AP_census_tract, aes(color = black_perc, fill = black_perc)) +
    geom_sf() + 
  labs(title = "Median bedroom rent in census tract")

```

```{r}
D_AP_census_tract %>%
ggplot( aes(y = n_active
            , x = med_bdr_rent)) + 
  geom_point(size=1,
        alpha=0.5) +
  labs ( y = "Number of active sheds permits in a census tract",
        x = "Median of bedroom rent in a census tract",
        title = "Area with higher rent per bedroom usually have more active sheds")
```

```{r}
ggplot(D_AP_census_tract, aes(y = n_active, x = pov_perc)) + 
  geom_point(size=1,
        alpha=0.5) +
  labs ( y = "Number of active sheds permits in census tract",
        x = "Percentage of people in poverty in census tract",
        title = "There are more active sheds permits in richer areas")
```

The reason could be:

 1. There are fewer buildings higher than 6-story in poor areas (no related data).

2.  The buildings condition in poor area are just too good to require sidewalk sheds.
3.  There are as many deteriorated building facades in poor area, but the expense is too high to erect a sidewalk sheds.

```{r}
ggplot(D_AP_census_tract, aes(y = n_active, x = age_before_2000)) + 
  geom_point(size=1,
        alpha=0.5) +
  labs ( y = "Number of active sheds permits in census tract",
        x = "Number of buildings built before 2000 in census tract",
        title = "With more older buildings, there more active sheds today.")
```

```{r}
ggplot(D_AP_census_tract, aes(y = n_active, x = black_tenure)) + 
  geom_point(size=1,
        alpha=0.5) +
  labs ( y = "Number of active sheds permits in census tract",
        x = "African American-only tenure in census tract",
        title = "Areas with more black-only tenure have fewer active sheds permits.")
```

### logistic regression

#### 1. For number of active sheds

Looks like a okay model

```{r}
D_mod_count<-glm(
  prevalent~
                    black_perc+pov_perc+summary_est+med_inc+age_before_2000+med_bdr_rent+per_tenure_size+med_value,
           family = "binomial",
           data = D_AP_census_tract )

summary(D_mod_count)
```


```{r fig.width=6, fig.height=6}
par(mfrow= c(2,2))
plot(D_mod_count)
```

#### For average duration of active sheds

Not a good model

```{r}
D_mod_long<-glm(
  long~
          black_perc+pov_perc+summary_est+med_inc+age_before_2000+med_bdr_rent+per_tenure_size+med_value,
           family = "binomial",
           data = D_AP_census_tract )

summary(D_mod_count)
```

```{r fig.width=6, fig.height=6}
par(mfrow= c(2,2))
plot(D_mod_long)
```

```{r}
cdists<-cooks.distance(D_mod_long)
plot(cdists, cex = .5, pch = 16, xlab = "Index", ylab = "Cook's distance")
influential_to_rm <- cdists %>% sort(decreasing = T) %>% head(10) %>% names %>% as.numeric
p_resid<-resid(D_mod_long, "pearson") %>% sort(decreasing = T)
large_resid<- as.numeric(names(which(p_resid>20)))
D_mod_long$data[c(large_resid,influential_to_rm),] %>% View
```

```{r}
pdata <- predict(D_mod_many, newdata = D_AP_census_tract, type = "response")
confusionMatrix(data = as.numeric(pdata>0.5) %>% as.factor, reference = D_AP_census_tract$long %>% as.numeric %>% {.-1} %>% as.factor)
```
